Composable Generalized Consensus

### 2015-05-15 paper/paper.tex
Working on the presentation of CGC

Generalized Lattice Agreement \cite{FalerioETAL12GeneralizedLatticeAgreement} allows more efficient implementations of SMR when
commands are partitioned into read-only commands and write-only commands.
Composable Generalized Consensus supports the same optimizations as Generalized Lattice Agreement.
Notably, read-only commands are treated specially so as to optimize their
executions (see \cref{sec:readonly}).

@@

Our specifications all contain by default a state variable $crashed$ and
 a $Crash\left( x \right)$ transition which is always enabled. A server $x$ faithfully
executes the algorithm that is assigned to it until its identifier is added to
the variable $crashed$ by a $Crash\left( x \right)$ transitions.  When the identifier of a
server is in the set $crashed$, all its actions are disabled.  In an execution,
a server that is never added to the set $crashed$ is said correct.

@@

Before formally introducing our framework, let us see how we can build a modular implementation similar to Fast Paxos by first designing a fast but fragile modular GC algorithm, then combining it unchanged with a more resilient module, obtaining an algorithm similar to Fast Paxos.

@@

\item The $Abort\left( s, h\right)$ transition, for a switcher $s$, is enabled when $aborts\left[ s \right] = \quo{none}$ and the switcher has received
  at least one $\aseq{\quo{stopped}, a, h}$ message. Its effect is to set the variable $aborts\left[ s \right]$ to the history $h$ received in the $\quo{stopped}$ message. The history $aborts'\left[ s \right]$ is then called an abort history. Because the switcher has received a $\aseq{\quo{stopped},a,h}$ message from the acceptor $a$, the acceptor $a$ no longer accepts new proposals, therefore no history greater than $h$ can ever be learned anymore. Therefore, an abort history can be seen as a safe starting point for a new round to resume execution.

@@

@@

Informally, the composition is correct because abort histories are always safe in the sense that any of the abort histories could be the next learned history.
However not all can be learned and the initialization mechanism of the rounds ensures that at most one is chosen.
In the next section we present Composable Generalized Consensus, a specification of the behavior of rounds at the level of proposals, learned histories, init histories, and abort histories. Our composition theorem ensures that any set of rounds individually satisfying Composable Generalized Consensus can be composed in any order to yield an algorithm satisfying Composable Generalized Consensus.

@@

We now consider implementing Generalized Consensus as a sequence of \emph{Composable Generalized Consensus rounds}, where each round makes progress
ordering requests or aborts its execution and passes the baton to the next round in the sequence. An algorithm composed of a sequence of CGC instances
is called a \emph{CGC algorithm}.

On top of the traditional roles of proposer, acceptor, and learner, CGC
introduces the new role of \emph{switcher} and a new component, the
\emph{scheduling policy}. In CGC, the assignment of the roles of proposer,
acceptor, learner, and switcher can change in each rounds, allowing for
reconfiguration. 
In normal circumstances, the proposers, acceptors,
and learners of a CGC instance work together to implement Generalized Consensus.
However, if the environment becomes unfavorable to a running CGC instance, the
scheduling policy can instruct the round to abort its execution and pass the
baton, using the switchers, to a new round.

A switcher of an aborting round determines, from the state of the
acceptors, an \emph{abort sequence} that it transfers to one or several
switchers of the new round.  Upon receiving an abort sequence, a switcher of the
new round initializes the acceptors of the new round, which can then
proceed implementing SMR, until the new round itself aborts and the
process is repeated. 
We assume that in the first round, which has no previous round, every 
switchers receives the empty abort sequence.

The scheduling policy determines when to abort a CGC instance and which new CGC
round should take over. The scheduling policy must inform the switchers about the
identity of the next round and must guarantee that all switchers receive the
same information.  In this paper we abstract over the scheduling policy, assuming
that a round that aborts passes the baton to a nondeterministically chosen new
round. A reliable scheduling-policy component can realistically be implemented
as a replicated state machine running on a large number of servers otherwise
dedicated to other tasks
\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}. We will see
 in \cref{sec:rounds} that, as in Vertical Paxos, a reliable scheduling policy component allows
tolerating $f$ faults with $f+1$ replicas, while maintaining the long-term
resilience of the service by replacing faulty servers upon round change.

@@

CGC specifies the allowed behaviors of the proposers, learners, and switchers
\emph{of a single round}. Similarly to
Generalized Consensus, we use the state variables $crashed$, $proposed$,
and, for every learner $l$, $learned\left[ l \right]$.  Additionally,
we use the state variables $inits$, containing the abort sequences
received from the previous round, and $aborts$, containing the abort
sequences transfered to the next round.
Two consecutive CGC instances are composed by letting the $aborts$ variable of the
first round be the $inits$ variable of the second round.

Initially, $crashed$, $proposed$, for every
learner $l$, $learned\left[ l \right]$, $inits$, and $aborts$ are all empty sets.
The next-state relation consists of five types of transitions. The
first three, $Crash\left( x \right)$, $Propose\left( l, c
\right)$, and $Learn\left( l, cs \right)$, for a learner, proposer, or
switcher $x$, for a learner $l$, and for a sequence of commands
$cs$, have the same effect as in Generalized Consensus.
CGC introduces two new transitions, $Init\left(
s, cs \right)$ and $Abort\left( s, cs \right)$, for a switcher $s$ and a sequence of
commands $cs$. The transition $Init\left( s, cs \right)$ adds $cs$ to the set $inits$, and models a switcher receiving the
abort sequence $cs$ from the preceding round.
The transition $Abort\left( s, cs \right)$ adds $cs$ to the set $aborts$ and models a switcher transferring the abort sequence $cs$ to a switcher of the next round.
All the transition are guarded by the following conditions:
\begin{compactitem}
    \item[\textbf{Agreement}:] The set $\aset{learned'\left[ l
      \right] : l \in Learners}$ is compatible.
    \item[\textbf{Validity}:] There is a history $h$ belonging to $inits$ such that every abort history or learned history 
        is of the form $h \star cs$, where $cs$ contains only commands
        of the set $proposed$.
    \item[\textbf{Initialization}:] If $inits$ is empty, then no
        sequence can be learned and no switcher can abort.
    \item[\textbf{Irrevocability}:] In the case of a $Learn\left( l,h
        \right)$ transition, the new learned history $h$ is an 
        extension of $learned\left[ l \right]$.
    \item[\textbf{Safe Abort}:] Every abort sequence is an extension in the 
        partial order $\leq$ of a learned history.
\end{compactitem}
Note that Agreement and Irrevocability are the same as in Generalized Consensus.

>>>>

%Generalized Lattice Agreement \cite{FalerioETAL12GeneralizedLatticeAgreement} allows more efficient implementations of SMR when
%commands are partitioned into read-only commands and write-only commands.
%Composable Generalized Consensus supports the same optimizations as Generalized Lattice Agreement.
%Notably, read-only commands are treated specially so as to optimize their
%executions (see \cref{sec:readonly}).

@@

@@

In this section we introduce our framework by example, building a modular Generalized Consensus algorithm inspired from Fast Paxos by first designing a fast but fragile modular algorithm, then combining it unchanged with a more resilient algorithm.

@@

\item The $Abort\left( s, h\right)$ transition, for a switcher $s$, is enabled when $aborts\left[ s \right] = \quo{none}$ and the switcher has received
  at least one $\aseq{\quo{stopped}, a, h}$ message. Its effect is to set the variable $aborts\left[ s \right]$ to the history $h$ received in the $\quo{stopped}$ message. The history $aborts'\left[ s \right]$ is then called an abort history.

@@

Crucially, the composition of any number of rounds implements Generalized Consensus.
To understand why, observe that in an $Abort\left( s,h \right)$ transition, the switcher $s$ has received a $\aseq{\quo{stopped},a,h}$ message from the acceptor $a$, which means that the acceptor $a$ no longer accepts new proposals and that $h$ is the current value of the $learned\left[ a \right]$ variable. Therefore no history greater than $h$ can ever be learned anymore. An abort history can thus be seen as a safe starting point for a new round to resume execution.

@@

Observe that a Simple-Leader round maintains a similar invariant, relating abort histories to learn histories, than the BroadCast-Decide algorithm:
the learned histories are necessarily smaller than the abort histories in the partial order on histories. In both types of rounds, an abort history is
a safe starting point for initializing the variables $accepted\left[ a \right]$ of the acceptors. Moreover, both types of rounds guarantee that only a
single history will be used as a basis for the learners to learn new histories.

In the next section we present Composable Generalized Consensus, a specification of the behavior of rounds at the level of proposals, learned histories, init histories, and abort histories, which allows making precise the informal reasoning above.
Our composition theorem ensures that any set of rounds individually satisfying Composable Generalized Consensus can be composed in any order to yield an algorithm satisfying Composable Generalized Consensus.
Therefore we can devise many other types of rounds and compose them unchanged with our two BroadCast-Decide and Simple-Leader rounds.

@@

In this section we present Composable Generalized Consensus, our specification for composable Generalized Consensus rounds.

As described in the preceding section, the servers can have any of the four roles of proposer, acceptor, learner, and switcher. 
The assignment of the roles of proposer,
acceptor, learner, and switcher can change in each rounds, allowing for
reconfiguration.

We assume that in the first round, which has no previous round, every 
switchers receives the empty abort sequence.

\begin{comment}
The scheduling policy determines when to abort a CGC instance and which new CGC
round should take over. The scheduling policy must inform the switchers about the
identity of the next round and must guarantee that all switchers receive the
same information.  In this paper we abstract over the scheduling policy, assuming
that a round that aborts passes the baton to a nondeterministically chosen new
round. A reliable scheduling-policy component can realistically be implemented
as a replicated state machine running on a large number of servers otherwise
dedicated to other tasks
\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}. We will see
 in \cref{sec:rounds} that, as in Vertical Paxos, a reliable scheduling policy component allows
tolerating $f$ faults with $f+1$ replicas, while maintaining the long-term
resilience of the service by replacing faulty servers upon round change.
\end{comment}

@@

CGC specifies the allowed behaviors of the proposers, learners, and switchers
\emph{of a single round}. Similarly to
Generalized Consensus, we use the state variables $crashed$, $proposed$,
and, for every learner $l$, $learned\left[ l \right]$.  Additionally,
we use the state variables $inits$, containing the abort sequences
received from the previous round, and $aborts$, containing the abort
sequences transfered to the next round.
Two consecutive CGC rounds are composed by letting the $aborts$ variable of the
first round be the $inits$ variable of the second round.

Initially, $crashed$, $proposed$, for every learner $l$, $learned\left[ l \right]$, $inits$, and $aborts$ are all empty sets. The next-state relation
consists of four types of transitions. The first two, $Propose\left( l, c \right)$, and $Learn\left( l, h \right)$, for a learner $l$, and for a
history $h$, have the same effect as in Generalized Consensus. CGC introduces two new transitions, $Init\left( s, h \right)$ and $Abort\left( s, h
\right)$, for a switcher $s$ and a history $h$. The transition $Init\left( s,h \right)$ adds $h$ to the set $inits$, and models a switcher receiving
the abort sequence $h$ from the preceding round. The transition $Abort\left( s, h \right)$ adds $cs$ to the set $aborts$ and models a switcher
transferring the abort sequence $cs$ to a switcher of the next round. All the transition are guarded by the following conditions:
\begin{compactitem}
    \item[\textbf{Agreement}:] The set $\aset{learned'\left[ l
      \right] : l \in Learners}$ is compatible.
    \item[\textbf{Validity}:] There is a history $h$ belonging to $inits$ such that every abort history or learned history 
        is of the form $h \star cs$, where $cs$ contains only commands
        of the set $proposed$.
    \item[\textbf{Initialization}:] If $inits$ is empty, then no
        sequence can be learned and no switcher can abort.
    \item[\textbf{Irrevocability}:] In the case of a $Learn\left( l,h
        \right)$ transition, the new learned history $h$ is an 
        extension of $learned\left[ l \right]$.
    \item[\textbf{Safe Abort}:] Every abort sequence is an extension in the 
        partial order $\leq$ of a learned history.
\end{compactitem}
Note that Agreement and Irrevocability are the same as in Generalized Consensus.
